---
title: "Using ez-zarr in R"
author: "Charlotte Soneson, Michael Stadler"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_document:
        toc: true
        toc_float: true
        theme: united
vignette: >
  %\VignetteIndexEntry{Using ez-zarr in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette describes how to use the `ezzarr` R package, which provides a 
thin R wrapper around the [`ez-zarr`](https://pypi.org/project/ez-zarr/) 
python package. `ezzarr` provides an easy 
way to set up and activate a `conda` environment with all necessary 
dependencies for `ez-zarr`, and once the environment is activated, the 
`reticulate` package can be used to access the functionality in `ez-zarr`, 
to import and interact with image and plate objects from OME-Zarr files in R. 

# Load the `ezzarr` package

```{r setup}
library(ezzarr)
```

# Check the version of ez-zarr

The `getEzZarrVersion()` function returns the version of `ez-zarr` which will 
be installed and used by the current version of `ezzarr`. Calling this function 
will also build the `conda` environment if it does not already exist. 

```{r}
getEzZarrVersion()
```

# Get the path to the environment

`ezzarr` provides a convenience function to get the path to the `ez-zarr` 
`conda` environment. This can be useful e.g. in order to activate the 
environment outside of R. 

```{r}
getEzZarrEnvPath()
```

# Download example data to temporary directory

In the remainder of this vignette, we will illustrate the use of `ezzarr` by
means of an example data set, representing a maximum intensity projection of 
one well from a high-content screening plate, processed using 
[Fractal](https://fractal-analytics-platform.github.io/). The same data set 
is used in the `ez-zarr` 
[python documentation](https://fmicompbio.github.io/ez_zarr/getting_started_Image/).
The data is available on Zenodo; here, we download and unpack it to a temporary 
directory. 

```{r}
td <- tempdir()
url <- "https://zenodo.org/records/10519143/files/20200812-CardiomyocyteDifferentiation14-Cycle1_mip.zarr.zip"
download.file(url, destfile = file.path(td, basename(url)))
unzip(file.path(td, basename(url)), 
      exdir = td)
```

Before going further, we save the path to the unpacked data set in the 
`zarrpath` variable, and explore the content of the directory. 

```{r}
zarrpath <- file.path(td, sub("\\.zip$", "", basename(url)))
fs::dir_tree(zarrpath, recurse = 3)
```

# Activate the environment

Next, we activate the `conda` environment that contains `ez-zarr` and all 
necessary dependencies, via the `enableEzZarr()` function. In addition to 
activating the `conda` environment, this function returns a list of imported
python modules, which can be used to access python functions. 

```{r}
## Activate environment
env <- enableEzZarr()

## List imported modules
names(env)

## Switch matplotlib backend (for non-interactive use)
env$plt$switch_backend("agg")
```

# Load an example image

Now that the environment is activated, we can import an image from an 
OME-Zarr fileset. We do this by creating an object of the 
[`Image`](https://fmicompbio.github.io/ez_zarr/ome_zarr/#src.ez_zarr.ome_zarr.Image) 
class, defined in `ez-zarr`. 

```{r}
img <- env$ez_zarr$ome_zarr$Image(path = file.path(zarrpath, "B/03/0"))

img

img$tree(level = 1)
```


# Get details about the image

`ez-zarr` contains many 
[methods](https://fmicompbio.github.io/ez_zarr/ome_zarr/#src.ez_zarr.ome_zarr.Image) 
for accessing various aspects of an `Image` object. Here, we illustrate some 
of these. 

```{r}
## Channel information
length(img$nchannels_image)
img$channels[[1]]

## Label names
img$label_names

## Scales for a given pyramid level
unlist(img$get_scale(pyramid_level = "2"))

## Path
img$get_path()

## Image name
img$name

## Table names
img$table_names
```

# Plot image

Plots in `ez-zarr` are implemented using `matplotlib.pyplot` and display
correctly when using the plot methods interactively from python
(e.g. `img.plot()`) or from R via `reticulate` (e.g. `img$plot()`).
A special case is the use of plot methods in non-interactive Rmarkdown or
Quarto documents. In these, the plots are only shown for `python` code chunks,
which is used in the examples below (note the `r.img` syntax to refer to the
`img` object in the R process).

## With default settings

```{python}
r.img.plot()
```

The corresponding R code (for interactive use) would look like:

```{r, eval = FALSE}
img$plot()
```

## Adding a scale bar, limiting the channel range

```{python}
r.img.plot(channels = [0],
           channel_colors = ["white"],
           channel_ranges = [[100, 600]],
           scalebar_micrometer = 150,
           scalebar_color = "yellow",
           scalebar_position = "topleft",
           scalebar_label = True,
           fig_width_inch = 5,
           fig_height_inch = 4,
           fig_dpi = 100,
           axis_style = "micrometer")
```

The corresponding R code for interactive is shown below. Please note that
you need to use specific R data structures to satisfy the expected arguments
types of the python function. For example, a python list (`[...]`) has to be
created using an R list (`list(...)`) and not a vector. For a list of
the conversions that reticulate uses between R and python data structures see
[reticulate type conversions](https://rstudio.github.io/reticulate/articles/calling_python.html#type-conversions).

```{r, eval=FALSE}
img$plot(channels = list(0L),
         channel_colors = list("white"),
         channel_ranges = list(c(100, 600)),
         scalebar_micrometer = 150L,
         scalebar_color = "yellow",
         scalebar_position = "topleft",
         scalebar_label = TRUE,
         fig_width_inch = 5,
         fig_height_inch = 4,
         fig_dpi = 100,
         axis_style = "micrometer")
```

## Zoom in

```{python}
r.img.plot(pyramid_level = "0",
           upper_left_yx = [130, 140],
           lower_right_yx = [300, 310],
           scalebar_micrometer = 30,
           scalebar_color = "magenta",
           fig_width_inch = 5,
           fig_height_inch = 5,
           fig_dpi = 100)
```

## Overlay nuclei segmentation

```{python}
r.img.plot(label_name = "nuclei", 
           pyramid_level = "0",
           upper_left_yx = [130, 140],
           lower_right_yx = [300, 310],
           scalebar_micrometer = 30,
           scalebar_color = "magenta",
           fig_width_inch = 5,
           fig_height_inch = 5,
           fig_dpi = 100)
```

# Import and plot entire plate

Above, we imported and visualized a single image. `ez-zarr` also contains 
classes to represent high-content screening plates or other, arbitrary, 
collections of images. Here, we use the `import_plate()` function to read the 
whole plate (note that only one well was imaged). 

```{r}
plt <- env$ez_zarr$ome_zarr$import_plate(zarrpath)
```

Again, we use a python code chunk for the plot to be included in our
compiled notebook (not needed for interactive use):

```{python}
r.plt.plot()
```


# Session info

```{r}
sessionInfo()
```

